---
title: "Midterm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(glue)
library(tidyverse)
library(rio)
library(janitor)
library(gapr)
```

# Part A. 

## 1. Use the function above to download all of the data from the 2014-15 to 2017-18 school years and bind it into a single data frame, using a single function (i.e., one line of code). Note, this may take a minute or two to run, depending on your Internet speed.

```{r}
download_file <- function(year) {
  link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
  rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}
d <- map_dfr(15:18, download_file, .id=NULL)
```

## Cleaning

### Filter for only student groups coded as "White" or "Hispanic/Latino".

```{r}
d <- d %>%  
  filter(`Student Group` %in% c("White","Hispanic/Latino"))
```

### Select variables related to the number of students in each of the levels (1:4), and not percentages or collapsed levels.

```{r}
d <- d %>% 
  select(-starts_with("Percent"),
         -`Number Proficient (Level 3 or 4)`, 
         -`Participation Rate`, 
         -`Number of Participants`,
         -`Number Proficient`, 
         -ends_with("ID"), 
         -Subject) %>% 
  clean_names()

head(d) 
```

### Pivot Longer & Drop `NA`s

```{r}
d <- pivot_longer(
  d, 
  cols = c(number_level_1:number_level_4),
  names_to = "level",
  values_to = "n")

d <- drop_na(d,n)
```

### Recode the "level" value

```{r}
d <- d %>% 
  mutate(
    level = 
      recode(
        level,
        number_level_1= "1", 
        number_level_2 ="2", 
        number_level_3 = "3", 
        number_level_4 ="4")
    )
```

## 2. Sum the n across grades and school for each district within each academic year, student group, and level. Calculate the number of schools (unique schools listed) for each district. Also move your data to a wider format where the student groups become new columns, filled by the district n. Remove districts that did not report numbers for one or both student groups (i.e., drop missing data from the new columns). 

```{r}

## we are confident there is a way to do this with tidyverse,
## but we are getting really unexpected results with count()
## so this is our solution

schools_per_district <- d %>% 
  group_by(district) %>% 
  summarize(n_school = length(unique(school)))
  
d <- d %>% 
  group_by(academic_year, district, student_group, level) %>%
  summarize(
    n = sum(n)
    ) %>%
  pivot_wider(names_from = student_group, values_from = n) %>% 
  clean_names() %>% 
  drop_na(hispanic_latino, white) %>% 
  left_join(schools_per_district, by = 'district')

head(d)
```

# Part B. 

## 1. Estimate the average achievement differences in terms of an effect size for every school district in the state that reported data on both student groups (i.e., using the data we created above), for each academic year. Conduct this analysis twice, once using group_by() %>% nest() map() and once using nest_by() %>% summarize().

```{r}
d[1:4, ] %>% # data source - just first four rows for first district
  estimate_v("white", "hispanic_latino") # columns for reference/focal groups
```

